---
title: "Clustering"
author: "Jeff B"
date: "April 4, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE)
library(FNN)
library(mvtnorm)
library(mclust)
library(cluster)
library(fpc)
library(boot)
library(tree)
library(MASS)
library(randomForest)
library(MLmetrics)
```

```{r, echo=FALSE}
admissionsData <- read.csv("Admission_Predict_Ver1.1.csv")
#summary (admissionsData)
attach(admissionsData)
```

## Clustering

We begin by computing the respective pairwise distances in our data, and plotting the output.

```{r, echo=FALSE}
dg<-daisy(admissionsData[,-1], metric="gower")
pdist <- cmdscale(d=dg)
plot(pdist)
```

We quickly see that two clear groups appear. We can isolate these two groups using hierarchical clustering with single-linkage chaining. 

```{r, , echo=FALSE, eval=FALSE}
set.seed(413)
km <- kmeans(pdist, centers = 2)
#plotcluster(pdist, km$cluster)
```

```{r, echo=FALSE}
hms <- hclust(na.omit(dg), method="single")
#plot(hms)
plot(pdist, col=cutree(hms,2))
#plot(pdist)
```

We can then use scatterplots to show the entirety of the data, while still keeping the groups intact, to see if we can determine which predictors most affect these clusters.


```{r, echo=FALSE}
pairs(admissionsData[,-1], col=cutree(hms,2))
```

We notice that, using the single linkage chaining from above, we can predict whether or not a student performs research almost perfectly.

So, by applying Gower's Distance on all predictors and using single-linkage chaining, we have two clear clusters directly coinciding with the presence of a research variable. This tells us that we should use Research as a response variable in models, in addition to Chance of Admit.

We can now perform analyses on the data to attempt to predict a candidate's Chance of Admission, as well as the presence of Research Experience.

